Interval Probability and Measurement for Fuzzy Quantum Theories

### 2016-09-24 prob.tex
prob.tex Clean up the macros...

% \makeatletter
% \newtheoremstyle{indented}
%   {3pt}% space before
%   {3pt}% space after
%   {\addtolength{\@totalleftmargin}{3.5em}
%    \addtolength{\linewidth}{-3.5em}
%    \parshape 1 3.5em \linewidth}% body font
%   {}% indent
%   {\bfseries}% header font
%   {.}% punctuation
%   {.5em}% after theorem header
%   {}% header specification (empty for default)
% \makeatother
% \theoremstyle{indented}
\theoremstyle{remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{thm}{Theorem}
\newtheorem{case}{Case}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\qevents}{\ensuremath{\mathcal{E}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\pr}[2]{\langle #1, #2 \rangle}
\newcommand{\proj}[1]{|#1 \rangle\langle #1 |}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\poss}{{\mbox{\wesa{possible}}}}
\newcommand{\imposs}{{\mbox{\wesa{impossible}}}}
\newcommand{\likely}{{\mbox{\wesa{likely}}}}
\newcommand{\unlikely}{{\mbox{\wesa{unlikely}}}}
\newcommand{\necess}{{\mbox{\wesa{certain}}}}
\newcommand{\overflow}{{\mbox{\wesa{overflow}}}}
\newcommand{\unknown}{{\mbox{\wesa{unknown}}}}

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\yutsung}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}}}
\newcommand{\amr}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{green}{Amr says: #1}\end{minipage}}}
\newcommand{\ffzd}[1]{{\mathbb{F}^{d\;*}_{#1}}}
\def\C{{\mathbb{C}}}
\newcommand{\ff}[1]{\mathbb{F}_{#1}}
\newcommand{\expect}[2]{ \langle #1 | #2 | #1 \rangle}
\newcommand{\Tr}{\mathop{\mathrm{Tr}}\nolimits}
\newcommand{\gerardo}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{OliveGreen}{Gerardo says: #1}\end{minipage}}}
\newcommand{\andy}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{blue}{Andy says: #1}\end{minipage}}}
\usepackage{tensor}
\newcommand{\rme}{\mathrm{e}}
\newcommand{\rmi}{\mathrm{i}}
\usepackage[braket, qm]{qcircuit} 
\newcommand{\xyR}[1]{\xymatrixrowsep={#1}}
\newcommand{\xyC}[1]{\xymatrixcolsep={#1}}

@@

If we compare the counterexample in example~\ref{ex:three-dimensional-three-value}
and \ref{ex:three-dimensional-4-value}, we found example~\ref{ex:three-dimensional-4-value}
is not as bad as example~\ref{ex:three-dimensional-three-value}.
At least, the states map to $\imposs$ are all in the same plane in
example~\ref{ex:three-dimensional-4-value}. Maybe interval-valued
probability measures will be closer and closer to real-valued probability
measures if the intervals are more and more precise?

@@

The following example verify the core of a quantum interval-valued
probability measure is nonempty even if there is no natural correspondence.
For example, if $\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
there is no natural way to map from $\frac{1}{2}$, as both $\frac{1}{2}\in\left[0,\frac{1}{2}\right]=\unlikely$
and $\frac{1}{2}\in\left[\frac{1}{2},1\right]=\likely$. However,
we still can find a quantum interval-valued probability measure corresponding
to a quantum probability measure as in the following example.

\begin{example}[One-qubit quantum interval-valued probability measure]\label{ex:One-qubit-quantum-interval}
Given a two dimensional Hilbert space, a vector $\ket{\psi}$ can
be normalized and written in $\left(\begin{array}{c}
\cos\theta\\
\rme^{\rmi\gamma}\sin\theta
\end{array}\right)$, where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. Consider
a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
defined by 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0}) & = & \imposs\textrm{ ,}\\
\bar{\mu}(\mathbb{1}) & = & \necess\textrm{ ,}\\
\bar{\mu}(\proj{\psi}) & = & \begin{cases}
\necess & \textrm{, if }\theta=0\textrm{ ;}\\
\likely & \textrm{, if }0<\theta<\frac{\pi}{4}\textrm{ ;}\\
\likely & \textrm{, if }\theta=\frac{\pi}{4}\textrm{ and }0\le\gamma<\pi\textrm{ ;}\\
\unlikely & \textrm{, if }\theta=\frac{\pi}{4}\textrm{ and }\pi\le\gamma<2\pi\textrm{ ;}\\
\unlikely & \textrm{, if }\frac{\pi}{4}<\theta<\frac{\pi}{2}\textrm{ ;}\\
\imposs & \textrm{, if }\theta=\frac{\pi}{2}\textrm{ .}
\end{cases}
\end{eqnarray*}
Let $\ket{\psi^{\perp}}$ be the state perpendicular to $\ket{\psi}$.
Then, the vector representation of $\ket{\psi^{\perp}}$ can be simplified
as 
\[
\left(\begin{array}{c}
-\rme^{-\rmi\gamma}\sin\theta\\
\cos\theta
\end{array}\right)=-\rme^{-\rmi\gamma}\left(\begin{array}{c}
\cos\left(\frac{\pi}{2}-\theta\right)\\
\rme^{\rmi\left(\pi+\gamma\right)}\sin\left(\frac{\pi}{2}-\theta\right)
\end{array}\right)
\]
so we have 
\[
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi}\right)=\bar{\mu}\left(\proj{\psi^{\perp}}\right)=1\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(\proj{\psi}\right)\textrm{ .}
\]
Hence, $\bar{\mu}$ is a quantum $\mathscr{I}$-interval-valued probability
measure. Notice that for all projection~$P$, $\mu_{\ket{0}}^{B}(P)\in\bar{\mu}(P)$
implies that $\bar{\mu}(P)$ is one of quantum $\mathscr{I}$-interval-valued
probability measures corresponding to $\mu_{\ket{0}}^{B}(P)$. However,
$\mu_{\ket{0}}^{B}(P)$ can correspond to infinity many quantum $\mathscr{I}$-interval-valued
probability measures by twisting the definition involving $\gamma$.\qed\end{example}

@@

\begin{example}[Three-dimensional quantum 4-value interval-valued
probability measure]\label{ex:three-dimensional-4-value} \footnote{\yutsung{Need to verify this example is convex!}}Given
a three dimensional Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Consider\emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$,
and a quantum $\mathscr{I}$-interval-valued probability measure~$\bar{\mu}$
defined as follow: 
\begin{enumerate}
\item $\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$
and $\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\necess$. 
\item Consider the states orthogonal to $\ket{0}$ and $\ket{\ps}$ respectively,
$\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
and $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\ms}+\cos\theta\ket{2}$,
where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. Half of
those states need special treatment, i.e., when where $0\le\theta<\frac{\pi}{2}$
and $0\le\gamma<\pi$, we define $\bar{\mu}(\proj{0_{\theta,\gamma}^{\perp}})=\ket{\ps_{\theta,\gamma}^{\perp}}=\likely$
and $\bar{\mu}(\mathbb{1}-\proj{0_{\theta,\gamma}^{\perp}})=\bar{\mu}(\mathbb{1}-\proj{\ps_{\theta,\gamma}^{\perp}})=\unlikely$. 
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$. 
\end{enumerate}
$\bar{\mu}$ cannot correspond to any quantum probability measure
because of the exact same reason as in example~\ref{ex:three-dimensional-three-value}.
Before we verify that $\bar{\mu}$ is a quantum interval-valued probability
measure, we need a notation for cross product. For $\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}+\alpha_{2}\ket{2}=\begin{pmatrix}\alpha_{0}\\
\alpha_{1}\\
\alpha_{2}
\end{pmatrix}$ and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}+\beta_{2}\ket{2}=\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\beta_{2}
\end{pmatrix}$, their cross product is defined as follow. 
\[
\ket{\psi\times\phi}=\ket{\psi}\times\ket{\phi}=\begin{vmatrix}\alpha_{1}^{*} & \beta_{1}^{*}\\
\alpha_{2}^{*} & \beta_{2}^{*}
\end{vmatrix}\ket{0}+\begin{vmatrix}\alpha_{2}^{*} & \beta_{2}^{*}\\
\alpha_{0}^{*} & \beta_{0}^{*}
\end{vmatrix}\ket{1}+\begin{vmatrix}\alpha_{0}^{*} & \beta_{0}^{*}\\
\alpha_{1}^{*} & \beta_{1}^{*}
\end{vmatrix}\ket{2}\textrm{ ,}
\]
where $\begin{vmatrix}\alpha & \beta\\
\alpha' & \beta'
\end{vmatrix}=\alpha\beta'-\alpha'\beta$. Notice that we put the complex conjugate in the cross product, so
that the result of cross product is orthogonal to $\ket{\psi}$ and
$\ket{\phi}$. 
\begin{eqnarray*}
\ip{\psi\times\phi}{\psi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\alpha_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\alpha_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\alpha_{2}=0\textrm{ ,}\\
\ip{\psi\times\phi}{\phi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\beta_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\beta_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\beta_{2}=0\textrm{ .}
\end{eqnarray*}
Also, even if $\ket{\psi}$ and $\ket{\phi}$ are normalized, their
cross product~$\ket{\psi\times\phi}$ need not be normalized as usual.

>>>>

% \makeatletter
% \newtheoremstyle{indented}
%   {3pt}% space before
%   {3pt}% space after
%   {\addtolength{\@totalleftmargin}{3.5em}
%    \addtolength{\linewidth}{-3.5em}
%    \parshape 1 3.5em \linewidth}% body font
%   {}% indent
%   {\bfseries}% header font
%   {.}% punctuation
%   {.5em}% after theorem header
%   {}% header specification (empty for default)
% \makeatother
% \theoremstyle{indented}
\theoremstyle{remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{thm}{Theorem}
\newtheorem{case}{Case}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\qevents}{\ensuremath{\mathcal{E}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\poss}{{\mbox{\wesa{possible}}}}
\newcommand{\imposs}{{\mbox{\wesa{impossible}}}}
\newcommand{\likely}{{\mbox{\wesa{likely}}}}
\newcommand{\unlikely}{{\mbox{\wesa{unlikely}}}}
\newcommand{\necess}{{\mbox{\wesa{certain}}}}
\newcommand{\overflow}{{\mbox{\wesa{overflow}}}}
\newcommand{\unknown}{{\mbox{\wesa{unknown}}}}

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\yutsung}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}}}
\newcommand{\amr}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{green}{Amr says: #1}\end{minipage}}}
\newcommand{\ffzd}[1]{{\mathbb{F}^{d\;*}_{#1}}}
\def\C{{\mathbb{C}}}
\newcommand{\ff}[1]{\mathbb{F}_{#1}}
\newcommand{\expect}[2]{ \langle #1 | #2 | #1 \rangle}
\newcommand{\Tr}{\mathop{\mathrm{Tr}}\nolimits}
\newcommand{\gerardo}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{OliveGreen}{Gerardo says: #1}\end{minipage}}}
\newcommand{\andy}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{blue}{Andy says: #1}\end{minipage}}}
\usepackage{tensor}
\newcommand{\rme}{\mathrm{e}}
\newcommand{\rmi}{\mathrm{i}}
\usepackage[braket]{qcircuit} 
\newcommand{\xyR}[1]{\xymatrixrowsep={#1}}
\newcommand{\xyC}[1]{\xymatrixcolsep={#1}}
\newcommand{\ip}[2]{\ensuremath{\left\langle{#1}\middle\vert{#2}\right\rangle}}
\newcommand{\melem}[3]{\ensuremath{\left\langle{#1}\middle\vert{#2}\middle\vert{#3}\right\rangle}}
\newcommand{\op}[2]{\ensuremath{\left\vert{#1}\middle\rangle\middle\langle{#2}\right\vert}}
\newcommand{\pr}[2]{\ip{#1}{#2}}
\newcommand{\proj}[1]{\op{#1}{#1}}

@@

If we compare the counterexample in example~\ref{ex:three-dimensional-three-value}
and \ref{ex:three-dimensional-4-value}, we find that example~\ref{ex:three-dimensional-three-value}
really use the property of $\unknown$ so that it cannot hold when
the intervals are more precise. Also, example~\ref{ex:three-dimensional-4-value}
is not as bad as example~\ref{ex:three-dimensional-three-value}
in a sense that the states map to $\imposs$ are all in the same plane
in example~\ref{ex:three-dimensional-4-value}. Maybe interval-valued
probability measures will be closer and closer to real-valued probability
measures if the intervals are more and more precise?

@@

@@

\begin{example}[Three-dimensional quantum 4-value interval-valued
probability measure]\label{ex:three-dimensional-4-value} \footnote{\yutsung{Need to verify this example is convex!}}Given
a three dimensional Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Consider\emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$,
and a quantum $\mathscr{I}$-interval-valued probability measure~$\bar{\mu}$
defined as follow: 
\begin{enumerate}
\item $\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$
and $\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\necess$. 
\item Consider the states orthogonal to $\ket{0}$ and $\ket{\ps}$ respectively,
$\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
and $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\ms}+\cos\theta\ket{2}$,
where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. Half of
those states need special treatment, i.e., when where $0\le\theta<\frac{\pi}{2}$
and $0\le\gamma<\pi$, we define $\bar{\mu}(\proj{0_{\theta,\gamma}^{\perp}})=\ket{\ps_{\theta,\gamma}^{\perp}}=\likely$
and $\bar{\mu}(\mathbb{1}-\proj{0_{\theta,\gamma}^{\perp}})=\bar{\mu}(\mathbb{1}-\proj{\ps_{\theta,\gamma}^{\perp}})=\unlikely$. 
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$. 
\end{enumerate}
Before we verify that $\bar{\mu}$ is a quantum interval-valued probability
measure, we need a notation for cross product. For $\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}+\alpha_{2}\ket{2}=\begin{pmatrix}\alpha_{0}\\
\alpha_{1}\\
\alpha_{2}
\end{pmatrix}$ and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}+\beta_{2}\ket{2}=\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\beta_{2}
\end{pmatrix}$, their cross product is defined as follow. 
\[
\ket{\psi\times\phi}=\ket{\psi}\times\ket{\phi}=\begin{vmatrix}\alpha_{1}^{*} & \beta_{1}^{*}\\
\alpha_{2}^{*} & \beta_{2}^{*}
\end{vmatrix}\ket{0}+\begin{vmatrix}\alpha_{2}^{*} & \beta_{2}^{*}\\
\alpha_{0}^{*} & \beta_{0}^{*}
\end{vmatrix}\ket{1}+\begin{vmatrix}\alpha_{0}^{*} & \beta_{0}^{*}\\
\alpha_{1}^{*} & \beta_{1}^{*}
\end{vmatrix}\ket{2}\textrm{ ,}
\]
where $\begin{vmatrix}\alpha & \beta\\
\alpha' & \beta'
\end{vmatrix}=\alpha\beta'-\alpha'\beta$. Notice that we put the complex conjugate in the cross product, so
that the result of cross product is orthogonal to $\ket{\psi}$ and
$\ket{\phi}$. 
\begin{eqnarray*}
\ip{\psi\times\phi}{\psi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\alpha_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\alpha_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\alpha_{2}=0\textrm{ ,}\\
\ip{\psi\times\phi}{\phi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\beta_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\beta_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\beta_{2}=0\textrm{ .}
\end{eqnarray*}
Also, even if $\ket{\psi}$ and $\ket{\phi}$ are normalized, their
cross product~$\ket{\psi\times\phi}$ need not be normalized as usual.
